---
title: Demonstation of cheng package
date: edited on 2016.04.12 
author: Cheng
output:
  html_document:
    css: custom.css
    fig_height: 4
    fig_width: 6.5
    highlight: zenburn
    theme: flatly
    toc: True
  pdf_document:
    fig_height: 3.5
    fig_width: 5
---
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Demonstration of cheng package}
-->

```{r setup, echo=FALSE, warning=FALSE, message = FALSE}
require(knitr)
require(cheng)
require(EEM)
require(pls)
#opts_chunk$set(fig.width = 6.5, fig.height = 4, warning = FALSE, message = FALSE)
```

# Package installation
The package is put on private repo on bitbucket. It can be installed using `install_bitbucket` funciton
in devtools package.

```{r install, eval = FALSE}
library(devtools)
install_bitbucket("chengvt/cheng", auth_user="keisoku", 
			password="kikuchi", dependencies = TRUE)
library(cheng)
```

# EEM-related functions

## drawSpec: drawSpectra
Select and draw excitation or emission spectra. This function should also go to EEM package in the future.

```{r drawSpec}
require(EEM)
data(applejuice)
country <- sapply(strsplit(names(applejuice), split = "-"), "[", 1)

# ggplot
drawSpec(unfold(applejuice), EX = 340)
drawSpec(unfold(applejuice), EX = 340, group = country)
drawSpec(unfold(applejuice), EM = 400, group = country)

# base plot
drawSpec(unfold(applejuice), EX = 340, group = country, ggplot = FALSE)
drawSpec(unfold(applejuice), EM = 400, group = country, ggplot = FALSE)

```

## delZeroCol
Useful for deleting columns which are filled with zero for autoscaling operation.
```{r delZeroCol}
data(applejuice)
applejuice_uf <- unfold(applejuice)
dim(applejuice_uf)
applejuice_uf_nozero <- delZeroCol(applejuice_uf)
dim(applejuice_uf_nozero)
```

# PLS-related functions (using pls package)

## getR2: getR2 value from `mvr` class object
Alternative to pls package's `R2`. While `R2` requires a `newdata` dataframe which combines
both predictors and target, `getR2` lets you put in `newx` and `newy` separately. 
Aside from that, declaring `estimate` in `getR2` remind you which value you got. 

```{r getR2}
require(pls)
data(yarn)
model <- plsr(density ~ NIR, 6, data = yarn, validation = "CV")
R2(model)
getR2(model, estimate = "train") # return R2 at particular ncomp without intercept value
getR2(model, estimate = "CV")

```

## getRMSE: getRMSE value from `mvr` class object
Alternative to pls package's `RMSEP`. While `RMSEP` requires a `newdata` dataframe which combines
both predictors and target, `getRMSE` lets you put in `newx` and `newy` separately. 
Aside from that, declaring `estimate` in `getRMSE` remind you which value you got. 

```{r getRMSE}
require(pls)
data(yarn)
model <- plsr(density ~ NIR, 6, data = yarn, validation = "CV")
RMSEP(model)
getRMSE(model, estimate = "train") # return RMSE at particular ncomp without intercept value
getRMSE(model, estimate = "CV")

```

## getVIP: getVIP from `mvr` class object

```{r getVIP}
# load data
require(EEM)
data(gluten)
gluten_uf <- unfold(gluten) # unfold list into matrix
# delete columns with NA values
index <- colSums(is.na(gluten_uf)) == 0
gluten_uf <- gluten_uf[, index]
gluten_ratio <- as.numeric(names(gluten))

# build pls model using pls model
require(pls)
model <- plsr(gluten_ratio ~ gluten_uf, ncomp = 10, method = "oscorespls")
drawEEM(getVIP(model), 1) # color contour
drawEEMgg(getVIP(model), 1, nlevel = 50) # normal contour
```

## plot_ncomp: Plotting selected ncomp
Plotting ncomp of pls model.

```{r plot_ncomp}
require(pls)
data(yarn)
model <- plsr(density ~ NIR, 15, data = yarn)

# plot ncomp
validationplot(model) # available function in pls package
plot_ncomp(model) # offer a new format for easier view
plot.new()
plot_ncomp(model, ncomp = 4) # fill in the point of the selected LV

# now build pls model again using cross-validation
model <- plsr(density ~ NIR, 15, data = yarn, validation = "CV")
plot_ncomp(model) # notice that y-axis label change to RMSECV
plot_ncomp(model, ncomp = 4) # fill in the point of the selected LV
```

## plotVIP
A lazy function which binds `drawEEM` and `getVIP` together.

```{r plotVIP}
# load data
require(EEM)
data(gluten)
gluten_uf <- unfold(gluten) # unfold list into matrix
# delete columns with NA values
index <- colSums(is.na(gluten_uf)) == 0
gluten_uf <- gluten_uf[, index]
gluten_ratio <- as.numeric(names(gluten))

# build pls model using pls model
require(pls)
model <- plsr(gluten_ratio ~ gluten_uf, ncomp = 10, method = "oscorespls")
plotVIP(model) # color contour
```

## plsplot: Plotting pls prediction result

```{r plsplot}
require(pls)
data(yarn)
model <- plsr(density ~ NIR, 6, data = yarn, validation = "CV")
plsplot(model) # calibration set result
plsplot(model, estimate = "CV") # cross validation set result

## customizing the graphs
plsplot(model, location = "topleft") # change legend position
plsplot(model, show = "R2") # show only R2
plsplot(model, show = "R2", round = 4) # round to four digits
plsplot(model, fitline = FALSE) # get rid of fitline
plsplot(model, show = "R2", cex.stats = 3) # bigger stats font
plsplot(model, cex.lab = 1.5, cex.main = 2) # bigger labels font
```

## plsplot2: Plotting pls prediction result
`plsplot2` plots both calibration and validation on the same graph
```{r plsplot2}
require(pls)
data(yarn)
yarn.cal <- yarn[yarn$train,]
yarn.val <- yarn[!yarn$train,]
model <- plsr(density ~ NIR, 15, data = yarn.cal, validation = "CV")
plot_ncomp(model)
ncomp <- 4 # 4 components seem to be appropriate
model <- plsr(density ~ NIR, ncomp, data = yarn.cal) # recalculate

plsplot(model) # calibration
plsplot(model, newx = yarn.val$NIR, newy = yarn.val$density) # validation

# now put those two plots together
plsplot2(model, newx = yarn.val$NIR, newy = yarn.val$density) # calibration and validation
plsplot2(model, newx = yarn.val$NIR, newy = yarn.val$density, col.cal = "forestgreen", col.val = "skyblue") # change point colors
```

## trainPLS
`trainPLS` can be used to train PLS for train dataset by cross-validation. The preprocessing method will be optimized automatically from mean-center, normalize + mean-center and autoscale. However, the number of latent variables has to be determined manually. 

```{r trainPLS, fig.width=9, fig.height=7}
# load data
require(EEM)
data(gluten)
gluten_uf <- unfold(gluten) # unfold list into matrix
# delete columns with NA values
index <- colSums(is.na(gluten_uf)) == 0
gluten_uf <- gluten_uf[, index]
gluten_ratio <- as.numeric(names(gluten))

# build pls model using pls model
require(cheng)
trainPLS(x = gluten_uf, y = gluten_ratio)
```

if the number of componenents is too high, you can try increasing the threshold, which is default at 0.02

```{r, fig.width=9, fig.height=7}
trainPLS(x = gluten_uf, y = gluten_ratio, threshold = 0.1)
```

## trainPLS2
Train PLS for train dataset by cross-validation. This is different from trainPLS as 
you have to specify the preprocessing method manually. In addition, the variable reduction by
VIP can be done automatically with number of cycles specifiable.
```{r trainPLS2, eval = FALSE}
result <- trainPLS2(x = gluten_uf, y = gluten_ratio, cycles = 2)
```
The number of variables will decrease with each cycle. 

# Raman-related functions
## addLabel
```{r}
data(raman)
plotRaman(raman)
# labeling lines at x=1300
addLabel(raman, 1300)
```

## plotRaman
Plot raman spectra

```{r}
data(raman)
plotRaman(raman)
plotRaman(raman, col = "black") # all black
plotRaman(raman, cex.lab = 1.5) # bigger axis label
plotRaman(raman, col = cm.colors(47)) # change color pallette 
 
# plot by group
cultivar <- getText(rownames(raman), 2)
plotRaman(raman, group = cultivar)
```

## plotRamanLoading

```{r}
data(raman)
PCA <- prcomp(raman)
plotRamanLoading(PCA, ncomp = 1:2)
```

## readRaman
Read raw Raman files

```{r, eval = FALSE}
data <- readRaman(getwd()) # data in working directory
data <- readRaman("folder") # data in folder
data <- readRaman("file.txt") # specify file
data[1:5,1:5]
```

# Misc
## biplot

Some customizations were added to `biplot` function of stats package.
The usages were illustrated below.

```{r biplot}
require(EEM)
data(applejuice)
applejuice_uf <- unfold(applejuice) # unfold list into matrix
# get country of apple production
country <- sapply(strsplit(names(applejuice), split = "-"), "[", 1)

# select peaks
local_peak <- findLocalMax(applejuice, n = 1)
index <- colnames(applejuice_uf) %in% local_peak
applejuice_uf_selectedPeak <- applejuice_uf[,index, drop = FALSE]

# PCA
result <- prcomp(applejuice_uf_selectedPeak)

# create color palette for x points
library(RColorBrewer)
xcol <- brewer.pal(3, "Dark2")

# biplot 2: color the scores by group
biplot2(result, xlab = prcompname(result ,1), ylab = prcompname(result,2), 
xlabs = country, xcol = xcol)

# biplot3: turn scores into points and color them by group
biplot3(result, xlab = prcompname(result ,1), ylab = prcompname(result,2), 
xlabs = country, xcol = xcol)

# biplot4: same as biplot3 but move legend outside
biplot4(result, xlab = prcompname(result ,1), ylab = prcompname(result,2), 
xlabs = country, xcol = xcol, legendinset = 0.07)

```

## getText

```{r getText}
string <- "country_cultivar_fruit_1"
getText(string, 2) # get the second group of string

# different separator
string <- "country~cultivar~fruit~1"
getText(string, 2, sep = "~") 
```

## plotScore_ellipse
Plotting ellipse for PCA score plots

```{r plotScore_ellipse}
data(applejuice)
applejuice_uf <- unfold(applejuice)
PCA <- prcomp(applejuice_uf)
plotScore_ellipse(PCA)

# change level
plotScore_ellipse(PCA, level = 0.8) 

# manually set x,y ranges
plotScore_ellipse(PCA, xlim = c(-9000, 9000), ylim = c(-4000, 4000)) 

# fill in circles
plotScore_ellipse(PCA, fill = TRUE, 
fill.alpha = 0.2, xlim = c(-9000, 9000), ylim = c(-4000, 4000))
```